Inefficient Concessions and Mediation

### 2016-12-03 presentation.tex
finish first draft of presentation

\section{Overview}
\subsection{Preview}
\begin{frame}
\frametitle{Unrecognized States}
Unrecognized states (URS): control and govern territory, seek recognition
\pause
\begin{itemize}[<+->]
  \item Six current URS more than 20 yrs old (+ Eastern Ukraine)
	\item No explanation in the literature for how this can be a stable outcome
\end{itemize}
\end{frame}

@@

\pause
We demonstrate unrecognized statehood can be a ``status quo'' outcome
\pause
\begin{itemize}[<+->]
	\item (SPN) Equilibrium in a repeated game
	\item Four players
		\begin{itemize}[<+->]
			\item Home government
			\item Secessionist elite
			\item Patron state
			\item International community
		\end{itemize}
	\item State variable: Status Quo (SQ) payoffs for secessionists
\end{itemize}

@@

\begin{frame}{The General Idea}

\pause
In each period: secessionists and gov't each choose (simultaneously) among $\left\{\text{Fight, Status Quo, Cede}\right\}$
\pause
		\begin{itemize}
			\item We need Status Quo to be a stage game best response for both secessionists and gov't
		\end{itemize}

\pause
\vskip.2in
We add some more realistic elements:
\pause
		\begin{enumerate}[<+->]
			\item Unrecognized status reduces Status Quo payoffs of secessionists each period
			\item Patron and int'l community can make investments in both actors' payoffs
		\end{enumerate}

@@

\begin{frame}{Outline of Talk}
\pause
\begin{enumerate}[<+->]
	\item M
	\item S
	\item W
\end{enumerate}
\end{frame}

\section{Model}
\subsection{Economic and Political Structure}
\begin{frame}{Timeline}
\pause
\end{frame}

\subsection{The Players}

\begin{frame}{Home State Actors}
	\pause
Central Government of Home State ($g$): desires reunification \\
\pause
Secessionists ($s$): desire recognized independence
\pause
\begin{itemize}	
	\item Central issue of contention: recognized independence vs. reunification.
\end{itemize}

\pause
\vskip.2in
Assumptions:
\begin{itemize}[<+->]
	\item Issue of status is indivisible, highly valued by both sides
	\item Insufficient credible side payments for easy settlement
\end{itemize}
\end{frame}

\begin{frame}{Stage Game between Home Gov't $\&$ Secessionists}
\pause

@@

\begin{frame}{Benchmark Model}
\pause
Assume two types: $\de_h$ and $\de_l$
\begin{itemize}[<+->]
	\item $\de_h > \de^* > \de_l$ where $\de^*$ is the cutoff for sustaining (Trust,Trust) eqm
	\item value of any concession to recipient = cost to giver
\end{itemize}

\pause
\vskip.2in
Some equilibria of interest
\begin{itemize}[<+->]
	\item Pool on `Fight'
	\item Separating without concessions
	\item Separating through concessions
\end{itemize}
\end{frame}

@@

	\pause
	\vskip.2in
	\begin{itemize}
		\item The smallest concession is $p(T+D)$
		\item If $p$ is low, high types are better off in the `fight' pooling equilibrium
	\end{itemize}
\end{frame}

\section{Inefficient Concessions}
\subsection{Add Money Burning}
\begin{frame}{S}
Assume cost of concession is $g$, the concession itself
\begin{itemize}
  \item Allow giver of concession to also choose $0 \leq e \leq 1$, where benefit of concession is $eg$
\end{itemize}

\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{P}
  With the

\end{beamerboxesrounded}

@@

\begin{frame}{S}
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{P}
  A
\end{beamerboxesrounded}

\section{Conclusion}
\subsection{}
\begin{frame}{Conclusion}
\begin{itemize}[<+->]
	\item We present 
\end{itemize}

>>>>

\section{Overview}
\subsection{Preview}
\begin{frame}
\frametitle{Inefficient Concessions and Mediation}

In conflict scenarios, concessions are sometimes inefficient 
\pause
\begin{itemize}[<+->]
  \item We show inefficient concessions may be preferable if efficient concessions have potential future cost
	\item Cost: concessions used against you by negotiating partner who violates agreement
	\item Can you trust that partner will make peace?
\end{itemize}

\pause
\vskip.2in
New explanation for mediation
\pause
\begin{itemize}[<+->]
  \item Can remove uncertainty about ability of negotiating partner to commit to peace
	\item Removes need for inefficient concessions
\end{itemize}
\end{frame}

@@

\pause
Start with simple, two-player repeated Prisoners' Dilemma
\pause
\begin{itemize}[<+->]
	\item Asymmetric information about $\de_i$: can partner commit to peace?
\end{itemize}

\pause
\vskip.2in
Add time zero: let partners give concessions with both signaling and material value
\pause
\begin{itemize}[<+->]
	\item Let partners destroy some/all of the material value
	\item Let material value provide help/harm to the giver
\end{itemize}

\pause
\vskip.2in
Mediator removes uncertainty about partner's $\de$

@@

\begin{frame}{Literature}
\pause
Signaling: Spence (1973), but signal has material value
\pause
		\begin{itemize}
			\item cost/benefit tradeoff differs b/c of $\de_i$, $C_i(g)$
		\end{itemize}

\pause
Gift-Giving: Camerer (1988), Prendergast $\&$ Stole (2001)
\pause
		\begin{itemize}
			\item Source of inefficiency differs: gifts can be used against giver
		\end{itemize}

\pause
Conflict: Slantchev (2011), Arena (2013)
\pause
		\begin{itemize}[<+->]
			\item Commitment to peace, not resolve to fight
			\item Concessions are costly signals instead of bargaining chip
			\item Costly signals are concessions instead of proof of resolve
		\end{itemize}

\pause
Mediation: Fey and Ramsay (2008, 2011), Horner et al. (2010)
\pause
		\begin{itemize}
			\item Information is about ability to commit, not resolve
		\end{itemize}

@@

\section{Model}
\subsection{}
\begin{frame}{Timeline}
\pause
\begin{enumerate}[<+->]
	\item[-1.] Nature independently determines types of Country 1 and Country 2: $\de_i \in \left\{\de_h,\de_l\right\}$
	\item[0.] Countries simultaneously give costly concessions: $g_1, g_2 \in \mathbb{R}_+$
	\item[1--$\infty$.] Countries engage in a simultaneous Prisoners' Dilemma interaction 
\end{enumerate}

\begin{frame}{Stage Game}
\pause

@@

\begin{frame}{Benchmark Model}
\pause
Assume two types: $\de_h$ and $\de_l$
\begin{itemize}[<+->]
	\item $\de_h > \de^* > \de_l$ where $\de^*$ is the cutoff for sustaining (Trust,Trust) eqm
	\item $p$: probability of high type
	\item Low type's cost to give concession $g$: $c_l(g)>c_h(g) =g$
\end{itemize}

\pause
\vskip.2in
Some equilibria of interest
\pause
\begin{itemize}[<+->]
	\item Pool on `Fight'
	\item Separating without concessions
	\item Separating through concessions
\end{itemize}
\end{frame}

@@

	\pause
	\vskip.2in
	\begin{itemize}[<+->]
		\item The smallest concession is $c_l^{-1}\left(p(T+D)\right)$
		\item If $p$ is low, high types are better off in the `fight' pooling equilibrium
	\end{itemize}
\end{frame}

\section{Inefficient Concessions}
\subsection{Add Money Burning}
\begin{frame}{Burning money is unattractive}
\pause
Allow giver of concession to also choose $0 \leq e \leq 1$, where benefit of concession is $eg$

\vskip.1in
\pause
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Theorem 3}
  When concessions confer no material help or harm on the giver, it is optimal to give efficient gifts ($e=1$) in a separating eqm.
\end{beamerboxesrounded}
\pause
\begin{itemize}[<+->]
	\item The benefit of the gift appears on both sides of the incentive constraint for both individuals, so cancels out
	\item The benefit appears in the high type's expected utility
\end{itemize}
\end{frame}

\begin{frame}[label=hurt]
\frametitle{Concessions can hurt the giver}
\pause
Now assume a low type will use concessions against you, i.e. to increase their payoffs from fighting \hyperlink{Supplemental}{\beamerbutton{Modified Payoffs}}
\vskip.1in
\pause
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Theorem 4}
  There are parameters under which the optimal equilibrium features inefficient concessions.
\end{beamerboxesrounded}
\[
  U_h=(1-\de_h)(-g) + \de_h \left[pT(1+eg)+peg+(1-p)(W-D(1+eg))\right]
\]
\pause

\begin{itemize}[<+->]
	\item If $p$ is low, concessions likely to be used against you, so remove material value
	\item Conjecture: Pooling on `Fight' optimal over larger set of parameters
\end{itemize}
\end{frame}

\section{Mediation}
\subsection{}
\begin{frame}{Mediator as mechanism designer}
\pause
`Manipulative' mediator: parties report their types, must deliver the stipulated concessions
\pause
\begin{itemize}[<+->]
  \item Mechanism: if two high types, give concession and play `Trust'; Otherwise, no concession and `Fight'
	\item Concession is necessary to get truthful revelation
	\item Cost of concession must be greater than benefit for low type
\end{itemize}
\vskip.1in
\pause
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Theorem 5}
  A mediator can eliminate inefficient concessions.
\end{beamerboxesrounded}

@@

\appendix
\section{Supplemental}
\begin{frame}[label=Supplemental,noframenumbering]
\begin{center}
Modified Stage Game Payoffs
\begin{tabular}{|l|c|c|}
  \hline      & Trust & Fight \\ \hline
	 Trust& \small$T(s_2+\alpha_2g_1)$, & \small$-D(m_2+(1-\alpha_2)g_1),$ \\
  & \small$T(s_1+\alpha_1g_2)$ & \small$T(s_1+\alpha_1g_2)  $\\ 
	&  & $+W(m_2+(1-\alpha_2)g_1) $\\ \hline
	Fight & \small$T(s_2+\alpha_2g_1)$& \small$W(m_1+(1-\alpha_1)g_2)$ \\
	      & \small$+W(m_1+(1-\alpha_1)g_2),$  & \small$-D(m_2+(1-\alpha_2)g_1),$ \\
	 & \small$-D(m_1+(1-\alpha_1)g_2) $ & \small$W(m_2+(1-\alpha_2)g_1)$ \\
	&  & \small$-D(m_1+(1-\alpha_1)g_2)$ \\ \hline
\end{tabular}
\end{center}

Back to \hyperlink{hurt}{\beamerbutton{Concessions can hurt the giver}}.
\end{frame}
